Effect of quality-adjusted production function inputs

### 2015-05-13 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Improved code for ESEE analysis.

<<Data, eval=TRUE>>=
dataPT <- filter(IST, Country == "PT")
dataUK <- filter(Leeds, Country == "UK")
@

<<Time Series Variables>>=
# This block of code sets up the time-series variables 
# to be used for each country.
# Adding this level of abstraction makes it easier 
# for us to change variables later.
# It also makes it easier to see what is going on 
# in each formula throughout the document.
# Use of substitute establishes an R "name"
# that can be a component of a formula.
# # Economic output
# y_PT <- substitute(iGDP)
# y_UK <- substitute(iGDP)
# # Portugal undajusted inputs
# k_PT_unadj <- substitute(iKstkS.L)
# l_PT_unadj <- substitute(iL)
# e_PT_unadj <- substitute(iXpMP)
# # Portugal quality-adjusted inputs
# k_PT_adj <- substitute(iKservS.L)
# l_PT_adj <- substitute(ihLest)
# e_PT_adj <- substitute(iUMP)
# # UK unadjusted inputs
# k_UK_unadj <- substitute(iKstkO.WwithRD)
# l_UK_unadj <- substitute(iL)
# e_UK_unadj <- substitute(iXp)
# # UK quality-adjusted inputs
# k_UK_adj <- substitute(iKservO.WwithRD)
# l_UK_adj <- substitute(ihLest)
# e_UK_adj <- substitute(iU)
# # Time variables
# t_PT <- substitute(iYear)
# t_UK <- substitute(iYear)

@@

y <- list()
k <- list()
l <- list()
e <- list()
t <- list()
# Economic output
y$PT <- substitute(iGDP)
y$UK <- substitute(iGDP)
# Portugal undajusted inputs
k$PT$unadj <- substitute(iKstkS.L)
l$PT$unadj <- substitute(iL)
e$PT$unadj <- substitute(iXpMP)
# Portugal quality-adjusted inputs
k$PT$adj <- substitute(iKservS.L)
l$PT$adj <- substitute(ihLest)
e$PT$adj <- substitute(iUMP)
# UK unadjusted inputs
k$UK$unadj <- substitute(iKstkO.WwithRD)
l$UK$unadj <- substitute(iL)
e$UK$unadj <- substitute(iXp)
# UK quality-adjusted inputs
k$UK$adj <- substitute(iKservO.WwithRD)
l$UK$adj <- substitute(ihLest)
e$UK$adj <- substitute(iU)
# Time variables
t$PT <- substitute(iYear)
t$UK <- substitute(iYear)

countries <- c(substitute(PT), substitute(UK))
adjusteds <- c(substitute(unadj), substitute(adj))
energies <- c(substitute(noE), substitute(withE))
f <- list()
for (country in c(substitute(PT), substitute(UK))){
  for (adjusted in adjusteds){
    for (energy in energies){
      if (energy == substitute(noE)){
        eval(substitute(f$ctry$adjst$ener <- y$ctry ~ k$ctry$adjst + l$ctry$adjst + t$ctry, 
                        list(ctry=country, adjst=adjusted, ener=energy)))
      } else if (energy == substitute(withE)){
        eval(substitute(f$ctry$adjst$ener <- y$ctry ~ k$ctry$adjst + l$ctry$adjst + e$ctry$adjst + t$ctry, 
                        list(ctry=country, adjst=adjusted, ener=energy)))
      } else {
        stop(paste("Unknown energy", energy, "in loop for f"))
      }
    }
  }
}

# Create some pre-econometric ratios
# Unadjusted variables
dataPT <- transform(dataPT, kPerGDP_unadj = eval(k_PT_unadj)/iGDP)
dataPT <- transform(dataPT, lPerGDP_unadj = eval(l_PT_unadj)/iGDP)
dataPT <- transform(dataPT, ePerGDP_unadj = eval(e_PT_unadj)/iGDP)
dataUK <- transform(dataUK, kPerGDP_unadj = eval(k_UK_unadj)/iGDP)
dataUK <- transform(dataUK, lPerGDP_unadj = eval(l_UK_unadj)/iGDP)
dataUK <- transform(dataUK, ePerGDP_unadj = eval(e_UK_unadj)/iGDP)
# Quality-adjusted variables
dataPT <- transform(dataPT, kPerGDP_adj = eval(k_PT_adj)/iGDP)
dataPT <- transform(dataPT, lPerGDP_adj = eval(l_PT_adj)/iGDP)
dataPT <- transform(dataPT, ePerGDP_adj = eval(e_PT_adj)/iGDP)
dataUK <- transform(dataUK, kPerGDP_adj = eval(k_UK_adj)/iGDP)
dataUK <- transform(dataUK, lPerGDP_adj = eval(l_UK_adj)/iGDP)
dataUK <- transform(dataUK, ePerGDP_adj = eval(e_UK_adj)/iGDP)

@@

<<Formulas>>=
# Set up the formulas for each type of analysis.
f <- NULL
f$PT$unadj$noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_PT, k = k_PT_unadj, 
                    l = l_PT_unadj, t = t_PT)))
f$PT$unadj$withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_PT, k = k_PT_unadj,
                    l = l_PT_unadj, e = e_PT_unadj, t = t_PT)))
f$PT$adj$noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_PT, k = k_PT_adj,
                    l = l_PT_adj, t = t_PT)))
f$PT$adj$withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_PT, k = k_PT_adj,
                    l = l_PT_adj, e = e_PT_adj, t = t_PT)))
f$UK$unadj$noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_UK, k = k_UK_unadj,
                    l = l_UK_unadj, t = t_UK)))
f$UK$unadj$withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_UK, k = k_UK_unadj,
                    l = l_UK_unadj, e = e_UK_unadj, t = t_UK)))
f$UK$adj$noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_UK, k = k_UK_adj,
                    l = l_UK_adj, t = t_UK)))
f$UK$adj$withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_UK, k = k_UK_adj,
                    l = l_UK_adj, e = e_UK_adj, t = t_UK)))
@

@@

\begin{table} \caption{Unadjusted time-series variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{y_PT}       & \Sexpr{y_UK}       \\
      Capital     & \Sexpr{k_PT_unadj} & \Sexpr{k_UK_unadj} \\
      Labor       & \Sexpr{l_PT_unadj} & \Sexpr{l_UK_unadj} \\
      Energy      & \Sexpr{e_PT_unadj} & \Sexpr{e_UK_unadj} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{Quality-adjusted time series variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{y_PT}     & \Sexpr{y_UK}     \\
      Capital     & \Sexpr{k_PT_adj} & \Sexpr{k_UK_adj} \\
      Labor       & \Sexpr{l_PT_adj} & \Sexpr{l_UK_adj} \\
      Energy      & \Sexpr{e_PT_adj} & \Sexpr{e_UK_adj} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<Historical Data, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
unadjVars_PT <- c(y_PT, k_PT_unadj, l_PT_unadj, e_PT_unadj)
unadjVars_UK <- c(y_UK, k_UK_unadj, l_UK_unadj, e_UK_unadj)
commonVars <- c("Year", "iYear", "Country")
unadjData_PT <- historicalData[ , as.character(unique(c(unadjVars_PT, commonVars)))] %>% filter(Country=="PT")
unadjData_UK <- historicalData[ , as.character(unique(c(unadjVars_UK, commonVars)))] %>% filter(Country=="UK")
unadjData <- plyr::rbind.fill(unadjData_PT, unadjData_UK)
unadjData$adjusted <- "Unadjusted"

adjVars_PT <- c(y_PT, k_PT_adj, l_PT_adj, e_PT_adj)
adjVars_UK <- c(y_UK, k_UK_adj, l_UK_adj, e_UK_adj)
adjData_PT <- historicalData[ , as.character(unique(c(adjVars_PT, commonVars)))] %>% filter(Country=="PT")
adjData_UK <- historicalData[ , as.character(unique(c(adjVars_UK, commonVars)))] %>% filter(Country=="UK")
adjData <- plyr::rbind.fill(adjData_PT, adjData_UK)
adjData$adjusted <- "Quality-adjusted"

@@

mdata <- melt(trimmedData, 
              measure.vars = unique(as.character(c(unadjVars_PT, unadjVars_UK, adjVars_PT, adjVars_UK))), 
              variable.name = "variable",
              value.name = "value")

# Relevel factors before plotting
mdata$Country <- relevelFactor(as.factor(mdata$Country), c("PT", "UK"))
mdata$adjusted <- relevelFactor(as.factor(mdata$adjusted), c("Unadjusted", "Quality-adjusted"))
mdata$variable <- relevelFactor(as.factor(mdata$variable), unique(as.character(
  c(y_PT, k_PT_unadj, l_PT_unadj, e_PT_unadj, k_PT_adj, l_PT_adj, e_PT_adj,
    y_UK, k_UK_unadj, l_UK_unadj, e_UK_unadj, k_UK_adj, l_UK_adj, e_UK_adj))))

historicalPlot(mdata, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ adjusted, 
               line_types = c(1,2,4,3,2,4,3,2,3,2,3)) + 
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  coord_cartesian(ylim=c(0.5, 13)) +
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  nest <- NULL
  nest$kl <- c(1,2)
  nest$kle <- c(1,2,3)
  nest$lek <- c(2,3,1)
  nest$ekl <- c(3,1,2)
  models <- NULL
  # Portugal, CD
  models$PT$unadj$noE$CD <- cdModel(formula = f$PT$unadj$noE, 
                                    data = dataPT, constrained = TRUE)
  models$PT$unadj$withE$CD <- cdModel(formula = f$PT$unadj$withE, 
                                      data = dataPT, constrained = TRUE)
  models$PT$adj$noE$CD <- cdModel(formula = f$PT$adj$noE, 
                                  data = dataPT, constrained = TRUE)
  models$PT$adj$withE$CD <- cdModel(formula = f$PT$adj$withE, 
                                    data = dataPT, constrained = TRUE)
  # UK, CD
  models$UK$unadj$noE$CD <- cdModel(formula = f$UK$unadj$noE, 
                                    data = dataUK, constrained = TRUE)
  models$UK$unadj$withE$CD <- cdModel(formula = f$UK$unadj$withE, 
                                      data = dataUK, constrained = TRUE)
  models$UK$adj$noE$CD <- cdModel(formula = f$UK$adj$noE, 
                                  data = dataUK, constrained = TRUE)
  models$UK$adj$withE$CD <- cdModel(formula = f$UK$adj$withE, 
                                    data = dataUK, constrained = TRUE)
  # PT, CESkl
  models$PT$unadj$noE$CESkl <- cesModel(formula = f$PT$unadj$noE, 
                                        nest = nest$kl, data = dataPT, constrained = TRUE)
  models$PT$adj$noE$CESkl <- cesModel(formula = f$PT$adj$noE, 
                                      nest = nest$kl, data = dataPT, constrained = TRUE)
  # PT, CESkle
  models$PT$unadj$withE$CESkle <- cesModel(formula = f$PT$unadj$withE, 
                                           nest = nest$kle, data = dataPT, constrained = TRUE)
  models$PT$adj$withE$CESkle <- cesModel(formula = f$PT$adj$withE, 
                                         nest = nest$kle, data = dataPT, constrained = TRUE)
  # PT, CESlek
  models$PT$unadj$withE$CESlek <- cesModel(formula = f$PT$unadj$withE, 
                                           nest = nest$lek, data = dataPT, constrained = TRUE)
  models$PT$adj$withE$CESlek <- cesModel(formula = f$PT$adj$withE, 
                                         nest = nest$lek, data = dataPT, constrained = TRUE)
  # PT, CESekl
  models$PT$unadj$withE$CESekl <- cesModel(formula = f$PT$unadj$withE, 
                                           nest = nest$ekl, data = dataPT, constrained = TRUE)
  models$PT$adj$withE$CESekl <- cesModel(formula = f$PT$adj$withE, 
                                         nest = nest$ekl, data = dataPT, constrained = TRUE)
  # UK, CESkl
  models$UK$unadj$noE$CESkl <- cesModel(formula = f$UK$unadj$noE, 
                                        nest = nest$kl, data = dataUK, constrained = TRUE)
  models$UK$adj$noE$CESkl <- cesModel(formula = f$UK$adj$noE, 
                                      nest = nest$kl, data = dataUK, constrained = TRUE)
  # UK, CESkle
  models$UK$unadj$withE$CESkle <- cesModel(formula = f$UK$unadj$withE, 
                                           nest = nest$kle, data = dataUK, constrained = TRUE)
  models$UK$adj$withE$CESkle <- cesModel(formula = f$UK$adj$withE, 
                                         nest = nest$kle, data = dataUK, constrained = TRUE)
  # UK, CESlek
  models$UK$unadj$withE$CESlek <- cesModel(formula = f$UK$unadj$withE, 
                                           nest = nest$lek, data = dataUK, constrained = TRUE)
  models$UK$adj$withE$CESlek <- cesModel(formula = f$UK$adj$withE, 
                                         nest = nest$lek, data = dataUK, constrained = TRUE)
  # UK, CESekl
  models$UK$unadj$withE$CESekl <- cesModel(formula = f$UK$unadj$withE, 
                                           nest = nest$ekl, data = dataUK, constrained = TRUE)
  models$UK$adj$withE$CESekl <- cesModel(formula = f$UK$adj$withE, 
                                         nest = nest$ekl, data = dataUK, constrained = TRUE)
  # Save models to disk.
  save(models, file=fileName)
}
@

>>>>

<<Data, eval=TRUE>>=
countries <- c("PT", "UK")
dataPT <- filter(IST, Country == "PT")
dataUK <- filter(Leeds, Country == "UK")
@

<<Time Series Variables>>=
# This block of code sets up the time-series variables 
# to be used for each country.
# Adding this level of abstraction makes it easier 
# for us to change variables later.
# It also makes it easier to see what is going on 
# in each formula throughout the document.
# Use of substitute establishes an R "name"
# that can be a component of a formula.

vars <-   
  list( 
    PT = list(
      response = list(unadjusted = "iGDP", adjusted = "iGDP"),
      capital = list(unadjusted ="iKstkS.L", adjusted = "iKservS.L"),
      labor = list(unadjusted = "iL", adjusted = "ihLest"),
      energy = list(unadjusted = "iXpMP", adjusted = "iUMP"),
      time = list(unadjusted = "iYear", adjusted = "iYear")
    ), 
    UK = list(
      response = list(unadjusted = "iGDP", adjusted = "iGDP"),
      capital = list(unadjusted = "iKstkO.WwithRD", adjusted = "iKservO.WwithRD"), 
      labor = list(unadjusted = "iL", adjusted = "ihLest"),
      energy = list(unadjusted = "iXp", adjusted = "iU"),
      time = list(unadjusted = "iYear", adjusted = "iYear")
    )
  )

# # Economic output
# y_PT <- substitute(iGDP)
# y_UK <- substitute(iGDP)
# # Portugal undajusted inputs
# k_PT_unadj <- substitute(iKstkS.L)
# l_PT_unadj <- substitute(iL)
# e_PT_unadj <- substitute(iXpMP)
# # Portugal quality-adjusted inputs
# k_PT_adj <- substitute(iKservS.L)
# l_PT_adj <- substitute(ihLest)
# e_PT_adj <- substitute(iUMP)
# # UK unadjusted inputs
# k_UK_unadj <- substitute(iKstkO.WwithRD)
# l_UK_unadj <- substitute(iL)
# e_UK_unadj <- substitute(iXp)
# # UK quality-adjusted inputs
# k_UK_adj <- substitute(iKservO.WwithRD)
# l_UK_adj <- substitute(ihLest)
# e_UK_adj <- substitute(iU)
# # Time variables
# t_PT <- substitute(iYear)
# t_UK <- substitute(iYear)

@@

# y <- list()
# k <- list()
# l <- list()
# e <- list()
# t <- list()
# # Economic output
# y$PT <- substitute(iGDP)
# y$UK <- substitute(iGDP)
# # Portugal undajusted inputs
# k$PT$unadj <- substitute(iKstkS.L)
# l$PT$unadj <- substitute(iL)
# e$PT$unadj <- substitute(iXpMP)
# # Portugal quality-adjusted inputs
# k$PT$adj <- substitute(iKservS.L)
# l$PT$adj <- substitute(ihLest)
# e$PT$adj <- substitute(iUMP)
# # UK unadjusted inputs
# k$UK$unadj <- substitute(iKstkO.WwithRD)
# l$UK$unadj <- substitute(iL)
# e$UK$unadj <- substitute(iXp)
# # UK quality-adjusted inputs
# k$UK$adj <- substitute(iKservO.WwithRD)
# l$UK$adj <- substitute(ihLest)
# e$UK$adj <- substitute(iU)
# # Time variables
# t$PT <- substitute(iYear)
# t$UK <- substitute(iYear)
# 
# foo <- function(country = c("UK", "PT"), adjusted=TRUE, energy=TRUE, envir=parent.frame()) {
#   country <- match.arg(country)
#   type <- if (adjusted) "adj" else "unadj"
#   res <- as.formula(
#       paste(y[[country]], 
#             paste(k[[country]][[type]], l[[country]][[type]], if (energy) e[[country]][[type]] else NULL, t[[country]], sep="+"), 
#             sep="~")
#     )
#   environment(res) <- envir
#   res
# }
# 
# 
# for (country in c("PT", "UK")) {
#   for (adjusted in c(TRUE, FALSE)) {
#     for (energy in c(TRUE, FALSE)) {
#       print(foo(country = country, adjusted = adjusted, energy=energy))
#     }
#   }
# }
#

# 
# countries <- c(substitute(PT), substitute(UK))
# adjusteds <- c(substitute(unadj), substitute(adj))
# energies <- c(substitute(noE), substitute(withE))
# f <- list()
# for (country in c(substitute(PT), substitute(UK))){
#   for (adjusted in adjusteds){
#     for (energy in energies){
#       if (energy == substitute(noE)){
#         eval(substitute(f$ctry$adjst$ener <- y$ctry ~ k$ctry$adjst + l$ctry$adjst + t$ctry, 
#                         list(ctry=country, adjst=adjusted, ener=energy)))
#       } else if (energy == substitute(withE)){
#         eval(substitute(f$ctry$adjst$ener <- y$ctry ~ k$ctry$adjst + l$ctry$adjst + e$ctry$adjst + t$ctry, 
#                         list(ctry=country, adjst=adjusted, ener=energy)))
#       } else {
#         stop(paste("Unknown energy", energy, "in loop for f"))
#       }
#     }
#   }
# }

# Create some pre-econometric ratios
# Unadjusted variables
dataPT <- transform(dataPT, kPerGDP.unadj = dataPT[[vars$PT$capital$unadjusted]]/dataPT[[vars$PT$response$unadjusted]])
dataPT <- transform(dataPT, kPerGDP.unadj = iKstkS.L/iGDP)
dataPT <- transform(dataPT, kPerGDP.unadj = as.name(vars$PT$capital$unadjusted)/iGDP)
dataPT <- transform(dataPT, kPerGDP.unadj = eval(substitute(as.name(vars$PT$capital$unadjusted)))/iGDP)
dataPT <- transform(dataPT, kPerGDP.unadj = eval(quote(as.name(vars$PT$capital$unadjusted)))/iGDP)

dataPT <- transform(dataPT, lPerGDP.unadj = eval(k_PT_unadj)/iGDP)
dataPT <- transform(dataPT, lPerGDP.unadj = eval(l_PT_unadj)/iGDP)
dataPT <- transform(dataPT, ePerGDP.unadj = eval(e_PT_unadj)/iGDP)
dataUK <- transform(dataUK, kPerGDP.unadj = eval(k_UK_unadj)/iGDP)
dataUK <- transform(dataUK, lPerGDP.unadj = eval(l_UK_unadj)/iGDP)
dataUK <- transform(dataUK, ePerGDP.unadj = eval(e_UK_unadj)/iGDP)
# Quality-adjusted variables
dataPT <- transform(dataPT, kPerGDP.adj = eval(k_PT_adj)/iGDP)
dataPT <- transform(dataPT, lPerGDP.adj = eval(l_PT_adj)/iGDP)
dataPT <- transform(dataPT, ePerGDP.adj = eval(e_PT_adj)/iGDP)
dataUK <- transform(dataUK, kPerGDP.adj = eval(k_UK_adj)/iGDP)
dataUK <- transform(dataUK, lPerGDP.adj = eval(l_UK_adj)/iGDP)
dataUK <- transform(dataUK, ePerGDP.adj = eval(e_UK_adj)/iGDP)

@@

<<Formulas>>=
# Set up the formulas for each type of analysis.
# f <- NULL
# f$PT$unadj$noE <- as.formula(
#   substitute( y ~ k + l + t,
#               list( y = y_PT, k = k_PT_unadj, 
#                     l = l_PT_unadj, t = t_PT)))
# f$PT$unadj$withE <- as.formula(
#   substitute( y ~ k + l + e + t,
#               list( y = y_PT, k = k_PT_unadj,
#                     l = l_PT_unadj, e = e_PT_unadj, t = t_PT)))
# f$PT$adj$noE <- as.formula(
#   substitute( y ~ k + l + t,
#               list( y = y_PT, k = k_PT_adj,
#                     l = l_PT_adj, t = t_PT)))
# f$PT$adj$withE <- as.formula(
#   substitute( y ~ k + l + e + t,
#               list( y = y_PT, k = k_PT_adj,
#                     l = l_PT_adj, e = e_PT_adj, t = t_PT)))
# f$UK$unadj$noE <- as.formula(
#   substitute( y ~ k + l + t,
#               list( y = y_UK, k = k_UK_unadj,
#                     l = l_UK_unadj, t = t_UK)))
# f$UK$unadj$withE <- as.formula(
#   substitute( y ~ k + l + e + t,
#               list( y = y_UK, k = k_UK_unadj,
#                     l = l_UK_unadj, e = e_UK_unadj, t = t_UK)))
# f$UK$adj$noE <- as.formula(
#   substitute( y ~ k + l + t,
#               list( y = y_UK, k = k_UK_adj,
#                     l = l_UK_adj, t = t_UK)))
# f$UK$adj$withE <- as.formula(
#   substitute( y ~ k + l + e + t,
#               list( y = y_UK, k = k_UK_adj,
#                     l = l_UK_adj, e = e_UK_adj, t = t_UK)))

makeFormula <- function(time.series, country, adjusted, energy, envir=parent.frame()) {
  country <- match.arg(country, names(time.series))
  if (! adjusted %in% c(FALSE, TRUE)){
    stop("adjusted must be a boolean")
  }
  if (! energy %in% c(FALSE, TRUE)){
    stop("energy must be a boolean")
  }
  type <- if (adjusted) "adjusted" else "unadjusted"
  res <- as.formula( 
    gsub( "\\+ *\\+", "+ ", 
          paste(time.series[[country]][["response"]][[type]], 
                paste(time.series[[country]][["capital"]][[type]], 
                      time.series[[country]][["labor"]][[type]], 
                      if (energy) {
                        time.series[[country]][["energy"]][[type]]
                      } else {
                        NULL
                      }, 
                      time.series[[country]][["time"]][[type]], sep="+"), 
                sep="~")
    )
  )
  environment(res) <- envir
  return(res)
}

f <- list()
for (country in countries) {
  for (adjusted in c(FALSE, TRUE)) {
    for (energy in c(FALSE, TRUE)) {
      newf <- makeFormula(vars, country = country, adjusted = adjusted, energy=energy)
      namef <- paste(country, if(adjusted) "adj" else "unadj", if(energy) "withE" else "noE", sep=".")
      newnames <- c(names(f), namef)
      f <- c(f, newf)
      names(f) <- newnames
    }
  }
}
@

@@

\begin{table} \caption{Unadjusted time-series variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{vars$PT$response$unadjusted}       & \Sexpr{vars$UK$response$unadjusted}       \\
      Capital     & \Sexpr{vars$PT$capital$unadjusted} & \Sexpr{vars$UK$capital$unadjusted} \\
      Labor       & \Sexpr{vars$PT$labor$unadjusted} & \Sexpr{vars$UK$labor$unadjusted} \\
      Energy      & \Sexpr{vars$PT$energy$unadjusted} & \Sexpr{vars$UK$energy$unadjusted} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{Quality-adjusted time series variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{vars$PT$response$adjusted}       & \Sexpr{vars$UK$response$adjusted}       \\
      Capital     & \Sexpr{vars$PT$capital$adjusted} & \Sexpr{vars$UK$capital$adjusted} \\
      Labor       & \Sexpr{vars$PT$labor$adjusted} & \Sexpr{vars$UK$labor$adjusted} \\
      Energy      & \Sexpr{vars$PT$energy$adjusted} & \Sexpr{vars$UK$energy$adjusted} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<Historical Data, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
unadjVars_PT <- c(vars$PT$response$unadjusted, 
                  vars$PT$capital$unadjusted, 
                  vars$PT$labor$unadjusted, 
                  vars$PT$energy$unadjusted)
unadjVars_UK <- c(vars$UK$response$unadjusted, 
                  vars$UK$capital$unadjusted, 
                  vars$UK$labor$unadjusted, 
                  vars$UK$energy$unadjusted)
commonVars <- c("Year", "iYear", "Country")
unadjData_PT <- historicalData[ , unique(c(unadjVars_PT, commonVars))] %>% filter(Country=="PT")
unadjData_UK <- historicalData[ , unique(c(unadjVars_UK, commonVars))] %>% filter(Country=="UK")
unadjData <- plyr::rbind.fill(unadjData_PT, unadjData_UK)
unadjData$type <- "Unadjusted"
adjVars_PT <- c(vars$PT$response$adjusted, 
                vars$PT$capital$adjusted, 
                vars$PT$labor$adjusted, 
                vars$PT$energy$adjusted)
adjVars_UK <- c(vars$UK$response$adjusted, 
                vars$UK$capital$adjusted, 
                vars$UK$labor$adjusted, 
                vars$UK$energy$adjusted)
adjData_PT <- historicalData[ , unique(c(adjVars_PT, commonVars))] %>% filter(Country=="PT")
adjData_UK <- historicalData[ , unique(c(adjVars_UK, commonVars))] %>% filter(Country=="UK")
adjData <- plyr::rbind.fill(adjData_PT, adjData_UK)
adjData$type <- "Quality-adjusted"

@@

mdata <- melt(trimmedData, 
              measure.vars = unique(c(unadjVars_PT, unadjVars_UK, adjVars_PT, adjVars_UK)), 
              variable.name = "variable",
              value.name = "value")

# Relevel factors before plotting
mdata$Country <- relevelFactor(as.factor(mdata$Country), c("PT", "UK"))
mdata$type <- relevelFactor(as.factor(mdata$type), c("Unadjusted", "Quality-adjusted"))
mdata$variable <- relevelFactor(as.factor(mdata$variable), unique(
  c(vars$PT$response$unadjusted, vars$PT$capital$unadjusted, vars$PT$labor$unadjusted, vars$PT$energy$unadjusted, 
    vars$PT$response$adjusted,   vars$PT$capital$adjusted,   vars$PT$labor$adjusted,   vars$PT$energy$adjusted, 
    vars$UK$response$unadjusted, vars$UK$capital$unadjusted, vars$UK$labor$unadjusted, vars$UK$energy$unadjusted, 
    vars$UK$response$adjusted,   vars$UK$capital$adjusted,   vars$UK$labor$adjusted,   vars$UK$energy$adjusted)))
historicalPlot(mdata, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ type, 
               line_types = c(1,2,4,3,2,4,3,2,3,2,3)) + 
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  coord_cartesian(ylim=c(0.5, 13)) +
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  nest <- NULL
  nest$kl <- c(1,2)
  nest$kle <- c(1,2,3)
  nest$lek <- c(2,3,1)
  nest$ekl <- c(3,1,2)

  models <- list()
  for ()

#   models <- NULL
#   # Portugal, CD
#   models$PT$unadj$noE$CD <- cdModel(formula = f$PT$unadj$noE, 
#                                     data = dataPT, constrained = TRUE)
#   models$PT$unadj$withE$CD <- cdModel(formula = f$PT$unadj$withE, 
#                                       data = dataPT, constrained = TRUE)
#   models$PT$adj$noE$CD <- cdModel(formula = f$PT$adj$noE, 
#                                   data = dataPT, constrained = TRUE)
#   models$PT$adj$withE$CD <- cdModel(formula = f$PT$adj$withE, 
#                                     data = dataPT, constrained = TRUE)
#   # UK, CD
#   models$UK$unadj$noE$CD <- cdModel(formula = f$UK$unadj$noE, 
#                                     data = dataUK, constrained = TRUE)
#   models$UK$unadj$withE$CD <- cdModel(formula = f$UK$unadj$withE, 
#                                       data = dataUK, constrained = TRUE)
#   models$UK$adj$noE$CD <- cdModel(formula = f$UK$adj$noE, 
#                                   data = dataUK, constrained = TRUE)
#   models$UK$adj$withE$CD <- cdModel(formula = f$UK$adj$withE, 
#                                     data = dataUK, constrained = TRUE)
#   # PT, CESkl
#   models$PT$unadj$noE$CESkl <- cesModel(formula = f$PT$unadj$noE, 
#                                         nest = nest$kl, data = dataPT, constrained = TRUE)
#   models$PT$adj$noE$CESkl <- cesModel(formula = f$PT$adj$noE, 
#                                       nest = nest$kl, data = dataPT, constrained = TRUE)
#   # PT, CESkle
#   models$PT$unadj$withE$CESkle <- cesModel(formula = f$PT$unadj$withE, 
#                                            nest = nest$kle, data = dataPT, constrained = TRUE)
#   models$PT$adj$withE$CESkle <- cesModel(formula = f$PT$adj$withE, 
#                                          nest = nest$kle, data = dataPT, constrained = TRUE)
#   # PT, CESlek
#   models$PT$unadj$withE$CESlek <- cesModel(formula = f$PT$unadj$withE, 
#                                            nest = nest$lek, data = dataPT, constrained = TRUE)
#   models$PT$adj$withE$CESlek <- cesModel(formula = f$PT$adj$withE, 
#                                          nest = nest$lek, data = dataPT, constrained = TRUE)
#   # PT, CESekl
#   models$PT$unadj$withE$CESekl <- cesModel(formula = f$PT$unadj$withE, 
#                                            nest = nest$ekl, data = dataPT, constrained = TRUE)
#   models$PT$adj$withE$CESekl <- cesModel(formula = f$PT$adj$withE, 
#                                          nest = nest$ekl, data = dataPT, constrained = TRUE)
#   # UK, CESkl
#   models$UK$unadj$noE$CESkl <- cesModel(formula = f$UK$unadj$noE, 
#                                         nest = nest$kl, data = dataUK, constrained = TRUE)
#   models$UK$adj$noE$CESkl <- cesModel(formula = f$UK$adj$noE, 
#                                       nest = nest$kl, data = dataUK, constrained = TRUE)
#   # UK, CESkle
#   models$UK$unadj$withE$CESkle <- cesModel(formula = f$UK$unadj$withE, 
#                                            nest = nest$kle, data = dataUK, constrained = TRUE)
#   models$UK$adj$withE$CESkle <- cesModel(formula = f$UK$adj$withE, 
#                                          nest = nest$kle, data = dataUK, constrained = TRUE)
#   # UK, CESlek
#   models$UK$unadj$withE$CESlek <- cesModel(formula = f$UK$unadj$withE, 
#                                            nest = nest$lek, data = dataUK, constrained = TRUE)
#   models$UK$adj$withE$CESlek <- cesModel(formula = f$UK$adj$withE, 
#                                          nest = nest$lek, data = dataUK, constrained = TRUE)
#   # UK, CESekl
#   models$UK$unadj$withE$CESekl <- cesModel(formula = f$UK$unadj$withE, 
#                                            nest = nest$ekl, data = dataUK, constrained = TRUE)
#   models$UK$adj$withE$CESekl <- cesModel(formula = f$UK$adj$withE, 
#                                          nest = nest$ekl, data = dataUK, constrained = TRUE)
  # Save models to disk.
  save(models, file=fileName)
}
@
